---
title: "R Notebook"
output: html_notebook
---
#Simulate the data
```{r}
#load required packages 
library(ggplot2)
library(readr)
library(rstatix)

#simulate the data

  #first specify the sample sizes 
ngroups = 2  #so i have two groups- multicultural vs. homogenous and then three within-participant conditions 
nsubsGroup = 60 #so 60 participants per groups 

sampleSizeRequired = 120 #a total of 120 participants needed
  
  #Next define the labels for the background groups and the conditions
background = c("Multicultural", "Homogenous") #here we are specifying the labels for the background conditions 

nConds = 3 #here i am specifying the number of conditions 
condLabels = c("Microinsult", "Microinvalid", "Normal") #the labels for the conditions

  #then we are going to create an empty data frame and specify the rows and the columns
df = data.frame(Participant=rep(0,sampleSizeRequired*nConds), 
                           Condition=rep("NA", sampleSizeRequired, nConds), #this fills in the condition column with NA 
                           background=rep("NA", sampleSizeRequired, nConds),#fills in the background column with NA
                           Likertresponse=rep(0, sampleSizeRequired, nConds), #fills in the likertresponse column with 0
                           stringAsFactors=TRUE
                           )

  #here we are specifying the values we want for the participants, so 1-120 but repeat it for each of the three conditions so you     have three rows for participant 1, three for participant 2 etc. 
df$Participant <- rep(1:sampleSizeRequired, each=nConds)  

  #here we are telling it to repeat the background characters - multicultural and homogenous for half of the 120 participants         becuase we want half of them to be multicultural and half homogenous. Then we want that to repeat again for each of the three       conditions. so that if a participant has 3 rows for each condition, each of those rows will say multicultural 
df$background <- rep(rep(background, each=sampleSizeRequired/2), each=nConds)

  #finally we need to fill in the condition column using the condition labels but we have to do it so that each participant gets one   #of the conditions so we use the time function
df$Condition <- rep(condLabels, times=sampleSizeRequired)

  #outcome measures: likert responses. we have to specify what the possible values could be and then replace the 0s in the likert     response column randomly with this function 
df$Likertresponse = sample(c(1,2,3,4,5), sampleSizeRequired, replace = TRUE)

#assume normality with the simulated data 

```

#Data Summary
First let's look at our data and inspect it to make sure it is in the proper format. We want the conditions and background to be in numeric format
```{r}
#this shows us means, median and sample size, also what type of data each variable is. 
summary(df)


#make our data for background and conditions numeric
back_num <- as.factor(df$background)
condition_num <- as.factor(df$Condition)


```
If we want to look at the entire dataset we can also do that. 
```{r}
df #shows you the entire dataset
```

Now we are going to look at the descriptive stats including mean and sd. 
```{r}
#we can also look at the descriptive stats for the data, first lets find the mean:
background_means <- aggregate(Likertresponse ~ back_num, data =  df, FUN = mean) #this gives us the means of the likert responses for the two background conditions
condition_means <- aggregate(Likertresponse ~ condition_num, data = df, FUN = mean)#this gives the means of the likert responses for the three within-group conditions

#we can also find the sd:
background_sd <- aggregate(Likertresponse ~ back_num, data =  df, FUN = sd) 
condition_sd <- aggregate(Likertresponse ~ condition_num, data = df, FUN = sd)



```


#statistical tests: 
Now we are going to analyze the data and look for significance using ANOVA and t-tests. 
```{r}
#we need to run a mixed 2*3 factorial ANOVA as there are both between subject factors and within

anova_results <- anova_test(
  data = df, 
  dv = Likertresponse,      # Dependent variable
  wid = Participant,        # Participant ID 
  between = background,     # Between-subjects factor
  within = Condition        # Within-subjects factor
)

print(anova_results)


```
The sphericity results show that the p<.05 therefore we can reject the null hypothesis meaning that the within subject effect of conditions has equal variances

The condition factor did show significant results. Meaning that participants show a perceived difference between the different conditions: microinsult, microinvalid, and normal interactions. 

#Post-hoc tests
Now we have to run a post-hoc test to see which of the three conditions significantly differed from each other: 
```{r}
post_hoc_results <- pairwise_t_test(
  df, 
  Likertresponse ~ Condition,  # Formula: dependent variable ~ factor | subject
  p.adjust.method = "bonferroni"  # Adjust p-values for multiple comparisons
)

# Print the results
print(post_hoc_results)

```
From these results we see that people rated the microinsults more severely than the normal interactions but there was no difference between microinsults and microinvalidations or microinvalidations and normal interactions. 


It is now important to run effect size measurements because we found a significant effect with the condition. We will use cohen's d.
```{r}


# Subset the data for the two conditions you're comparing
condition1_data <- df$Likertresponse[df$Condition == "Microinsult"]
condition2_data <- df$Likertresponse[df$Condition == "Normal"]

# Calculate the means for both conditions
mean_condition1 <- mean(condition1_data)
mean_condition2 <- mean(condition2_data)

# Calculate the standard deviations for both conditions
sd_condition1 <- sd(condition1_data)
sd_condition2 <- sd(condition2_data)

# Calculate the pooled standard deviation
pooled_sd <- sqrt((sd_condition1^2 + sd_condition2^2) / 2)

# Calculate Cohen's d
cohen_d <- (mean_condition1 - mean_condition2) / pooled_sd

# Print the result
print(cohen_d)

```
From this we get an effect size of .349 which is a between a small (.2) and medium effect (.5) meaning the difference between the microinsult and normal conditions was there but it isn't very large.

#Visualize the data 
Now we can visualize the data:
```{r}

#this plot shows interaction effects, we did not see any significant effects for interaction 
ggplot(df, aes(x = factor(condition_num), y = Likertresponse, color = factor(back_num), group = factor(back_num))) +
  stat_summary(fun = mean, geom = "point", size = 3) +  # Mean points
  stat_summary(fun = mean, geom = "line") +  # Line connecting means
  labs(title = "Interaction Effect of Background and Condition on Likert Responses",
       x = "Condition",
       y = "Mean Likert Response",
       color = "Background") +
  theme_minimal()

```


```{r}
#this shows boxplots for the distributions of each condiiton 
ggplot(df, aes(x = interaction(condition_num, back_num), y = Likertresponse, fill = back_num)) +
  geom_boxplot() +
  labs(x = "Condition", y = "Likert Response", fill = "Background") +
  scale_x_discrete(labels = c("Microinsult", "Microinvalid", "Normal",
                             "Microinsult", "Microinvalid", "Normal")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +  # Rotate x-axis labels
  ggtitle("Boxplot of Likert Responses by Condition and Background")

```
```{r}
#this is a double bar graph showing the means for the conditions, separated by background
ggplot(df, aes(x = factor(condition_num), y = Likertresponse, fill = factor(back_num))) +
  stat_summary(fun = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_cl_boot, geom = "errorbar", 
               position = position_dodge(width = 0.9), width = 0.2) +
  labs(title = "Mean Likert Responses by Condition & Background",
       x = "Condition",
       y = "Mean Likert Response",
       fill = "Background") +
  scale_fill_brewer(palette = "Set1") +  # Color scheme
  theme_minimal()
```
```{r}
#this scatterplot just shows all the individual responses and the means for each condition as well separated by background and condition
ggplot(df, aes(x = factor(condition_num), y = Likertresponse, color = factor(back_num))) +
  geom_jitter(width = 0.2, alpha = 0.5) +  # Add some jitter to avoid overplotting
  stat_summary(fun = mean, geom = "point", size = 3, shape = 17) +  # Show means
  stat_smooth(method = "lm", aes(group = back_num), se = FALSE) +  # Linear trend lines
  labs(title = "Scatter Plot of Likert Responses by Condition & Background",
       x = "Condition",
       y = "Likert Response",
       color = "Background") +
  theme_minimal()

```

